How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985

### 2022-09-14 Draft_September_2022/Full_draft_20220901.rmd
#finalminorcorrections

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem.* Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it represents the Amazon as a particular problem. These specific problems touch on common understandings of the Amazon as part of the larger socio-cultural history of the country, for example, in relation to issues of national sovereignty, economic integration, social development, or environmental conservation.

Different governments throughout Brazilian history have been described as proponents of a specific view of the Amazon [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. We conceptualize the Amazon region, forest, and people to be a policy object, this is a specific issue that deserves dedicated policy attention. While the military dictatorship(1964-85) is associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney's (1985-1990) and Lula's (2003-2010)presidencies are often tied to environmental conservation views. These monolithic representations advance that specific governments understand and act towards the Amazon as one problem, hiding within government diversity of both policies and perspectives. Albeit the current calls to understand the environment as a social-cultural construction and to identify the effect of culture on environmental outcomes [@lepolaindewaroux2021], we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Building on @hirschman1963 conceptualization of chosen problems and @bacchi2009 theory of problem-representation in policy, we propose a framework for identifying problem-constructions in discourse and investigate how the Brazilian Amazon has been constructed as a problem in transnational political discourses. We understand transnationalism[^1] as relationships that transcend nation-states, from the local level to the supranational level encompassing non-state actors [@keck1998].

@@

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices [@assuncao2015; @capobianco2019; @capobianco2021], are a crucial factor in decisions to deforest on the ground [@campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different settings, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

@@

Our empirical findings are threefold. First, we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (Climate Summits). Second, the same government adopts a diverse range of problem-constructions over time. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, briefly surpassing economic integration as more common problem-constructions from 2011 to 2016. Constructing the Amazon as an issue of sovereignty has increasingly become more pertinent after 2010. Finally, using logit models, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation as presidents move farther away from the Amazon region. That is, presidents talk to the economic and social needs of the people domestically while boasting about environmental policy internationally. Conceptually, this contributes to understanding the social construction of the Amazon in discourses and how these relates to policies. Empirically, we provide the first comprehensive overview of when, where, and how the Amazon is constructed as problems in presidential speeches.

@@

In "Journey towards Progress", @hirschman1963 draws a conceptual distinction between pressing problems (pressured by outside parties to the government) and chosen problems (chosen by the government at their discretion). Pressing problems can be either privileged or neglected, depending on the degree of pressure exercised by the interested groups. Chosen problems are those that governments select at their discretion. Problems can change from pressing to chosen across time and in space as a function of (a) solutions becoming available, (b) a change in the level of government control in society, or (c) a shift of interests from top policymakers [@hirschman1963]. @hirschman1975 (favourite) example of policymakers choosing a problem is President's Kubitschek decision of building Brasilia.

@@

We rely on the Amazonian historiography literature to identify possible problem-constructions and their connection to the socio-cultural history of the country. We refer to Amazonian historiography as the body of research conducted by social and environmental scientists that tell the story of diverse governmental programs that approach the Amazon region, forest, and its people as a policy object demanding dedicated action. From this literature, we identify three historical-political-social constructions: national sovereignty, economic integration, and environmental conservation. We also argue that a fourth one is missing, social development, that focuses on the rights and dignity of the inhabitants of the Amazonian region.

@@

The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon. In 1966, the Brazilian Military launched 'Operation Amazon', a policy to modernize the region based on three assumptions @acker2021. First, it assumes that nature should be conquered by mankind Second, it assumes that exploiting natural resources would render the Amazon region economically profitable. Third, it assumes that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructural projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. @capobianco2019 and @becker2005 describe the period from the 1950s to the 1980s in similar vein, as one of often centralized economic policies by the federal government.The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

@@

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. A common explanation for the creation of these institutions in the Amazonian literature is the impression of the lack of control over the market engendered years of centralized economic integration in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. @hochstetler2007 argue that during preparations for the summit, a new form of Brazilian environmentalism emerged: socio-environmentalism. They define it as an emphasis on the local livelihoods of traditional populations while protecting nature. @capobianco2019 argues that socio-environmentalism informed a series of policies throughout the 1990s and early 2000s towards preserving the forest, such as the establishment of several conservation units in the Amazon in 2001. The environmental conservation problem-constructions emphasize that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

@@

To analyze how presidents construct the Amazon as a problem, we build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil. Then, we proceed to identify all speeches that refer to the Amazon as a region, people, or forest in these speeches. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 946 speeches are, at least partially, about the Amazon out of the 6130.

Using the poldis R package [@sposito2021], we proceed to extract two sentences before and two sentences after the sentence in which the stem "amazon" appears. We opt for picking two sentences before and two sentences after, rather than words because sentences usually contain a cohesive idea. By doing so we create our unit of analysis: an Amazon statement. We use Amazon statements as our unit of analysis for two main reasons. First, working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. Second, it increases the number of observations and their variety meaningfully, allowing for more specificity in the analysis. This process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified. When an Amazonian statement contains two or more matches of the stem "amazon", we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

@@

Second, with the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazon Statements identified. We chose to hand.code half of the observations because since there are several nuances in discourse in how presidents talk about the Amazon over time and as a problem, we conservatively coded a relatively large training set for the subsequent automated machine learning. As well, this allows a robust validation set to verify the models. As the size of the training set should increase with the number of categories [@grimmer2022], we deemed four categories and 50% of the statements as reliable. Also, automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. The intercoder agreement for hand-coded categories was 85%, on average. For each non-matching coded observation, the co-authors discussed and sorted their disagreements.

@@

To analyze our data, we first present a series of different plots on the proportions of Amazonian statements and problem-constructions over time and by presidents. To test whether problem-constructions depend on where presidents speak, we run four separate logistic regressions. Each model takes the share of a specific problem-construction versus all other problem-constructions as their dependent variable. This means we try to predict the probability of a president constructing the Amazon as an issue of (1) environmental conservation, (2) economic integration, (3) social development, and (4) national sovereignty, separately, and in relation to the remaining three problem-constructions. Our main independent variable is the distance, in kilometers, between the state capital (within Brazil) or the country capital (outside of Brazil) to Manaus (the capital of the Amazonas State and centrally located in the region). In our model, we also control for annual deforestation rates, annual inflation rates, and election years in the model. We interpret the plots and model considering multiple Amazon-related events and policies over the last 30 years, with the support of existing literature, and their correlations with different presidents and locations.

@@

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15% [^2] of their speeches. The frequency at which the Amazon appears across presidential speeches inside, outside, and overall varies similarly across time (see Figure 7 in appendix).

@@

We observe five local maxima both inside and outside Brazil: 1989, 1992, 2009, 2015, and 2019. These points coincide with internal and external events that help us explain the rises and falls of the Amazon in presidential discourse. From 1985 to 1988, the mentions of the Amazon in official speeches increased from 4% to 14%. During this period, the latest Brazilian Constitution, which gave indigenous and traditional populations space to advocate for constitutional environmental rights and the protection of their territories [@hecht1990], was written. In 1989, however, the Amazon appeared in 32% of all speeches. This maximum correlates with the brutal murder of Chico Mendes in the last days of 1988. The incident caught unprecedented international attention and Sarney responded to this with a set of policies to address deforestation [@capobianco2021], including accepting to host the 1992 Earth Summit [@keck1998].

@@

From 2010 to 2018, except for 2015, we see a general decrease in mentions of the Amazon in all official presidential speeches. From 2010 to 2014, mentions of the Amazon went down from 18% to 8%. Disagreements related to the priority of environmental preservation over economic development, most notably in the case of approval of the New Forest Code in 2012, which regularized land ownership of many illegally deforested areas, were frequent. The period was also marked by political and economic instability which culminated in the impeachment of Rousseff in 2016. In 2015, the Amazon appeared in 19% of all speeches. This maximum coincides with the Paris summit which became a key turn in climate politics after the failures of Copenhagen. Brazil went to the Paris Summit with deforestation numbers slightly higher than Copenhagen and a perception that there was a turn towards less conservation inside the country. By 2016 mentions of Amazon in official speeches went down to 6%, the lowest share since 1985.

@@

The trends identified are especially interesting considering Amazonian policies and deforestation rates portrayed in Figure 3, above. First, deforestation rates have a strong negative correlation with environmental conservation. When deforestation rates increased from 1988 to 2004, and from 2012 until 2021, environmental conservation constructions decreased. As deforestation rates decreased from 2004 to 2012, environmental conservation constructions in presidential speeches increased. @capobianco2021 argues that the unprecedented decrease in deforestation we observed from 2004 to 2012 was a product of an increase in the perception of stronger federal policies and police presence in the Amazon region which, in turn, engendered a perception of a higher risk of being caught and fined for deforestation. The shift from constructing the Amazon as an issue of economic integration to an issue of environmental conservation generates a perception that illegal deforestation will be increasingly monitored and sanctioned. Second and relatedly, deforestation rates have a positive and strong correlation with economic integration. When deforestation increased from 1988 to 2004 and from 2012 to 2021, so did economic integration. When deforestation decreased from 2004 to 2012, economic integration decreased as well. This suggests that presidents might boast policy outcomes related to economic integration as justifications for deforestation.

@@

Table 1, below, shows a statistically significant relationship between constructing the Amazon as an issue of environmental conservation and distance from the Amazon. That is, the farthest away the president is from the Amazon region, the more likely the president is to construct it as an issue of environmental conservation. The opposite is true for both economic integration and social development, with statistically significant negative relationships. When speaking within the Amazon, presidents focus on social development and economic integration. This entails that conservation might be a desirable construction when speaking internationally about the Amazon, but not at the local level, where presidents highlight economic and social development. This multi-level game implies that presidents talk to the economic and social needs of the people domestically while boasting about environmentalism internationally.

>>>>

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem.* Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it represents the Amazon as a particular problem: national sovereignty, economic integration, social development, and environmental conservation, respectively.These specific problems touch on common understandings of the Amazon as part of the larger socio-cultural history of the country.

Different governments throughout Brazilian history have been described as proponents of a specific view of the Amazon [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. We conceptualize the Amazon region, forest, and people to be a policy object, this is a specific issue that deserves dedicated policy attention. While the military dictatorship(1964-85) is associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney's (1985-1990) and Lula's (2003-2010)presidencies are often tied to environmental conservation views. These monolithic representations advance the view that specific governments understand and act towards the Amazon as one problem, hiding within government diversity of both policies and perspectives. Albeit the current calls to understand the environment as a social-cultural construction and to identify the effect of culture on environmental outcomes [@lepolaindewaroux2021], we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Building on @hirschman1963 conceptualization of chosen problems and @bacchi2009 theory of problem-representation in policy, we propose a framework for identifying problem-constructions in discourse and investigate how the Brazilian Amazon has been constructed as a problem in transnational political discourses. We understand transnationalism[^1] as relationships that transcend nation-states, from the local level to the supranational level encompassing non-state actors [@keck1998].

@@

Although problem-construction takes place in a series of instances, we analyze the case of speeches by Brazilian presidents since 1985. Scholarly research about environmental policy in Brazil has focused on the role of the environmental bureaucracy [@silva-muller2022a], non-governmental organizations [@keck1998], legislation [@soares-filho2014], and international markets [@assuncao2015], and many others. Nevertheless, the role of the president remains understudied. We opt for presidential speeches for two reasons. First, presidential discourses have the power to introduce and justify the public policy, as well as shape its perception to broad audiences [@zarefsky2004; @gillion2016]. It legitimizes ways of thinking about the Amazon. In turn, policy perception is key for policy adoption and implementation [@alesina2009; @lopez2020]. This is especially pertinent to deforestation in Brazil as expectations of local actors, generated from material and discursive governmental practices, are a crucial factor in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021, @campbell2015]. Second, we argue that presidents can deploy problem-constructions that build policy objects as different problems depending on whom they are speaking to. Presidential discourses take place in different settings, from launching a new bridge in a small municipality in the middle of the Amazon to a speech at the UN general assembly in New York. Working with presidential discourses allows us to identify this variation in meaningful ways and better understand how the Amazon is discursively constructed in transnational politics.

@@

Our empirical findings are threefold. First, we find that the frequency of the Amazon as a topic in speeches inside and outside Brazil is equally driven by domestic events (e.g. the assassination of Chico Mendes) as well as international events (e.g. Climate Summits). Second, the same government adopts a diverse range of problem-constructions over time. While economic integration dominated discourse from 1985 to the mid-2000s, environmental conservation and social development constructions steadily grew in the 2000s, briefly surpassing economic integration as more common problem-constructions from 2011 to 2016. Constructing the Amazon as an issue of sovereignty has increasingly become more pertinent after 2010. Finally, using logit models, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation as presidents move farther away from the Amazon region. That is, presidents talk to the economic and social needs of the people domestically while boasting about environmental policy internationally. Conceptually, this contributes to understanding the social construction of the Amazon in discourses and how these relates to policies. Empirically, we provide the first comprehensive overview of when, where, and how the Amazon is constructed as problems in presidential speeches.

@@

In "Journey towards Progress", @hirschman1963 draws a conceptual distinction between pressing problems (pressured by outside parties to the government) and chosen problems (chosen by the government at their discretion). Pressing problems can be either privileged or neglected, depending on the degree of pressure exercised by the interested groups. Chosen problems are those that governments select at their discretion. Problems can change from pressing to chosen as a function of (a) solutions becoming available, (b) a change in the level of government control in society, or (c) a shift of interests from top policymakers [@hirschman1963]. @hirschman1975 's example of policymakers choosing a problem is President's Kubitschek decision of building Brasilia., which was not particularly pressed by any interest group.

@@

We rely on the Amazonian historiography literature to identify possible problem-constructions and their connection to the socio-cultural history of the country. We refer to Amazonian historiography as the body of research conducted by social and environmental scientists that tell the story of diverse governmental programs that approach the Amazon region, forest, and its people as a policy object demanding dedicated action. From this literature, we identify three historical-political constructions: national sovereignty, economic integration, and environmental conservation. We also argue that a fourth one is missing, social development, that focuses on the rights and dignity of the inhabitants of the Amazonian region.

@@

The Vargas dictatorship (1937-46) and the military dictatorship (1964-89) took over the task of modernizing the Amazon. In 1966, the Brazilian Military launched 'Operation Amazon', a policy to modernize the region based on three assumptions[@acker2021] . First, it assumes that nature should be conquered by mankind Second, it assumes that exploiting natural resources would render the Amazon region economically profitable. Third, it assumes that populating the region was necessary to integrate it into the country and exert control over the territory. Concretely, this meant a series of major infrastructural projects, incentives for settlers to expand the agricultural frontier, and the establishment of a tax-free zone in Manaus to attract industry. @capobianco2019 and @becker2005 describe the period from the 1950s to the 1980s in similar vein, as one of often centralized economic policies by the federal government.The economic integration problem-construction advances the view that the Brazilian Amazon needs to be developed and modernized. These policy solutions have at their core the development of the necessary infrastructure (physical, fiscal, or monetary) to integrate the region into the national and international economy.

@@

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. A common explanation for the creation of these institutions in the Amazonian literature is the impression of the lack of control over the market, engendered by years of centralized economic integration in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. @hochstetler2007 argue that during preparations for the summit, a new form of Brazilian environmentalism emerged: socio-environmentalism. They define it as an emphasis on the local livelihoods of traditional populations while protecting nature. @capobianco2019 argues that socio-environmentalism informed a series of policies throughout the 1990s and early 2000s towards preserving the forest, such as the establishment of several conservation units in the Amazon in 2001. The environmental conservation problem-constructions emphasize that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

@@

To analyze how presidents construct the Amazon as a problem, we build upon the dataset provided by [@cezar2020] which contains all official speeches by Brazilian Presidents from 1985 to 2019 scrapped from the archives of the Brazilian Presidential Library. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches for all the presidents of Brazil. Then, we proceed to identify all speeches that refer to the Amazon as a region, people, or forest. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 946 speeches are, at least partially, about the Amazon out of the 6130.

Using the poldis R package [@sposito2021], we proceed to extract two sentences before and two sentences after the sentence in which the stem "amazon" appears. We opt for picking two sentences around, rather than words because sentences usually contain a cohesive idea. By doing so we create our unit of analysis: an Amazon statement. We use Amazon statements as our unit of analysis for two main reasons. First, working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. Second, it increases the number of observations and their variety meaningfully, allowing for more specificity in the analysis. This process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified. When an Amazonian statement contains two or more matches of the stem "amazon", we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

@@

Second, with the codebook in hand, each one of the authors, separately, hand-coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazon Statements identified. We chose to hand code half of the observations because  there are several nuances in discourse in how presidents talk about the Amazon over time and as a problem.As the size of the training set should increase with the number of categories [@grimmer2022], we deemed four categories and 50% of the statements as reliable. This allows a robust validation set to verify the models.  Also, automating the coding of half of the observations saved the authors over one month of work in comparison to manual coding. The intercoder agreement for hand-coded categories was 85%, on average. For each non-matching coded observation, the co-authors discussed and sorted their disagreements.

@@

To analyze our data, we first present a series of different plots on the proportions of Amazonian statements and problem-constructions over time and by presidents. To test whether problem-constructions depend on where presidents speak, we run four separate logistic regressions. Each model takes the share of a specific problem-construction versus all other problem-constructions as their dependent variable. This means we try to predict the probability of a president constructing the Amazon as an issue of (1) environmental conservation, (2) economic integration, (3) social development, and (4) national sovereignty, separately, and in relation to the remaining three problem-constructions. Our main independent variable is the distance (in kilometers) from  the state or country capital where the speech took place to Manaus (the capital of the Amazonas State and centrally located in the region). In our model, we also control for annual deforestation rates, annual inflation rates, and election years in the model. We interpret the plots and model considering multiple Amazon-related events and policies over the last 30 years, with the support of existing literature, and their correlations with different presidents and locations.

@@

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15.4% [^2] of their speeches. The frequency at which the Amazon appears across presidential speeches inside, outside, and overall varies similarly across time (see Figure 7 in appendix).

@@

We observe five local maxima both inside and outside Brazil: 1989, 1992, 2009, 2015, and 2019. These points coincide with internal and external events that help us explain the rises and falls of the Amazon in presidential discourse. From 1985 to 1988, the mentions of the Amazon in official speeches increased from 4% to 14%. During this period, the latest Brazilian Constitution was written, which gave indigenous and traditional populations space to advocate for constitutional environmental rights and the protection of their territories [@hecht1990]. In 1989, however, the Amazon appeared in 32% of all speeches. This maximum correlates with the brutal murder of Chico Mendes in the last days of 1988. The incident caught unprecedented international attention and Sarney responded to this with a set of policies to address deforestation [@capobianco2021], including accepting to host the 1992 Earth Summit [@keck1998].

@@

From 2010 to 2018, except for 2015, we see a general decrease in mentions of the Amazon in all official presidential speeches. From 2010 to 2014, mentions of the Amazon went down from 18% to 8%. Disagreements related to the priority of environmental preservation over economic development, most notably in the case of approval of the New Forest Code in 2012, were frequent. The 2012 Forest Code regularized land ownership of many illegally deforested areas. The period was also marked by political and economic instability which culminated in the impeachment of Rousseff in 2016. In 2015, the Amazon appeared in 19% of all speeches. This maximum coincides with the Paris summit which became a key turn in climate politics after the failures of Copenhagen. Brazil went to the Paris Summit with deforestation numbers slightly higher than Copenhagen and a perception that there was a turn towards less conservation inside the country. By 2016 mentions of Amazon in official speeches went down to 6%, the lowest share since 1985.

@@

The trends identified are especially interesting considering Amazonian policies and deforestation rates portrayed in Figure 3, above. First, deforestation rates have a strong negative correlation with environmental conservation. When deforestation rates increased from 1988 to 2004, and from 2012 until 2021, environmental conservation constructions decreased. As deforestation rates decreased from 2004 to 2012, environmental conservation constructions in presidential speeches increased. @capobianco2021 argues that the unprecedented decrease in deforestation we observed from 2004 to 2012 was a product of an increase in the perception of stronger federal policies and police presence in the Amazon region which, in turn, generated a perception of a higher risk of being caught and fined for deforestation. The shift from constructing the Amazon as an issue of economic integration to an issue of environmental conservation generates a perception that illegal deforestation will be increasingly monitored and sanctioned. Second and relatedly, deforestation rates have a positive and strong correlation with economic integration. When deforestation increased from 1988 to 2004 and from 2012 to 2021, so did economic integration. When deforestation decreased from 2004 to 2012, economic integration decreased as well. This suggests that presidents might boast policy outcomes related to economic integration as justifications for deforestation.

@@

Table 1, below, shows a statistically significant relationship between constructing the Amazon as an issue of environmental conservation and distance from the Amazon. That is, the farthest away the president is from the Amazon region, the more likely the president is to construct it as an issue of environmental conservation. Second, the opposite is true for both economic integration and social development, with statistically significant negative relationships. When speaking within the Amazon, presidents focus on social development and economic integration. This entails that conservation might be a desirable construction when speaking internationally about the Amazon, but not at the local level, where presidents highlight economic and social development. This multi-level game implies that presidents talk to the economic and social needs of the people domestically while boasting about environmentalism internationally.
